
```{r}
sessionInfo()
```


```{r}
?install.packages()
```


```{r}
#| eval: FALSE
update.packages()
```


```{r}
install.packages("dplyr")
```


```{r}
?library()
```


```{r}
library(dplyr)
```


```{r}
install.packages("BiocManager")
```

```{r}
library(BiocManager)
```


```{r}
?install()
```


```{r}
install("PeacoQC")
```


```{r}
library(PeacoQC)
```

```{r}
sessionInfo()
```


```{r}
installed.packages()
```


```{r}
install.packages("remotes")
```


```{r}

```


```{r}
library(remotes)
```


```{r}
install_github("hally166/flowSpectrum")
```


```{r}
install.packages("purrr")
```


```{r}
force = TRUE
install("flowCore", force = TRUE)
```


```{r}
force = TRUE
```


```{r}
install.packages("stringr")
```


```{r}
install("ggcyto")
```


```{r}
install.packages("tidyr")
```


```{r}
install("openCyto")
```


```{r}
install("FlowSOM")
```


```{r}
install.packages("xml2")
```


```{r}
install_github("saeyslab/CytoNorm")
```


```{r}
install.packages("lubridate")
```


```{r}
install.packages("httr2")
```


```{r}
install("flowGate")
```


```{r}
install.packages("devtools")
```


```{r}
install("plotly")
```


```{r}
install_github("biosurf/cyCombine")
```


```{r}
install("CytoML")
```


```{r}
install.packages("Rtsne")
```


```{r}
install.packages("uwot")
```


```{r}
?PeacoQC
```

Homework:
1. PeacoQC determines peaks on channels in the flowframe. It then removes anomalies caused by clogs or flow rate changes etc.

2. bioconductor has 69 current packages. Mike Jiang has the most packages maintained/developed. 

Flowcore, flowViz, FlowSOM, ggcyto, CATALYST, flow density, flowcut, flowCHIC,flowGate, flowPlots: these packages seem to be the ones most useful in data visualisation, gating, and cleanup

3.
```{r}
install.packages("pak")
```


```{r}
?pak
```


```{r}
library(pak)
```


```{r}
pkg_install("broom")
```


```{r}
pkg_install("cytoMEM")
```

```{r}
pak::pkg_install("DillonHammill/CytoExploreR")
```

```{r}
packageVersion("superheat")
```


```{r}
install.packages("superheat")
```


```{r}
library(pak)
```


```{r}
pkg_install("superheat")
```