# Week9 Homework

:::{.callout-tip title="Problem 1"}
Modify one of the simpler functions (SecondFunction or similar), provide your own argument names, and modify the `message()`, `paste0` or `print()` functions to print a text style output. Generate a small vector of values, and iterate through your vector using one of the approaches we used.
:::

```{r}
MySecondFunction <- function(FastestCytometer){
    message("The Fastest Cytometer is the ", FastestCytometer)
}

MySecondFunction(FastestCytometer = "Attune")

Cytometers <- c("Attune", "Aurora", "FACSDiscover", "ID7000", "Xenith")
str(Cytometers)

MySecondFunction(FastestCytometer = Cytometers[1])

for (Name in Cytometers){
    MySecondFunction(FastestCytometer=Name)
}

```


:::{.callout-tip title="Problem 2"}
Using the initial framework for the CellConcentration function, retrieve several other keywords that are of interest, and incorporate them into the returned data.frame row. 
:::

```{r}
library(flowWorkspace)
library(flowGate)
library(ggplot2)
library(ggcyto)
library(dplyr)
library(purrr)

# StorageLocation <- file.path("course", "09_Functions", "data") # Interactive path
StorageLocation <- file.path("data") # Quarto Rendering path

fcs_files <- list.files(StorageLocation, pattern=".fcs", full.names=TRUE)

MyCytoset <- load_cytoset_from_fcs(fcs_files, transformation=FALSE, truncate_max_range=FALSE)
MyGatingSet <- GatingSet(MyCytoset)

MainDetectorsOnly <- c("V3-A", "B4-A", "R1-A")

Biexponential <- flowjo_biexp_trans(channelRange=4096, maxValue=262144,
     pos=4.5, neg=2, widthBasis=-500)
MyBiexTransform <- transformerList(MainDetectorsOnly, Biexponential)
transform(MyGatingSet, MyBiexTransform)

ggcyto(MyGatingSet[[1]], subset="root",
 aes(x="V3-A", y="R1-A")) + geom_hex(bins=100)

ggcyto(MyGatingSet[[1]], subset="root",
 aes(x="V3-A", y="B4-A")) + geom_hex(bins=100)

ggcyto(MyGatingSet[[1]], subset="root",
 aes(x="R1-A", y="B4-A")) + geom_hex(bins=100)

GatingTable <- tibble::tribble(
  ~filterId,    ~dims,                          ~subset,           
  "CD45",       list("B4-A", "SSC-A"),            "root",     
  "CD19",       list("V3-A", "SSC-A"),            "CD45",   
  "CD14",       list("R1-A", "SSC-A"),            "CD45", 
  "TandNKcells",  list("V3-A", "SSC-A"),            "CD45",
)

flowGate::gs_apply_gating_strategy(MyGatingSet, gating_strategy = GatingTable)
plot(MyGatingSet)

pData(MyGatingSet)$name

purrr::map(.x=MyGatingSet, .f=CellConcentration)

CellConcentration <- function(x){
    keyword(x)
}

KeywordList <- purrr::map(.x=MyGatingSet[1], .f=CellConcentration)
head(KeywordList[[1]],10000)

CellConcentration <- function(x, subset, dilutionMultiplier){
    Date <- keyword(x)$`$DATE` # Day-Month-Year
    InstrumentSerial <- keyword(x)$`$CYTSN`
    TotalFileEvents <- keyword(x)$`$TOT`
    Specimen <- keyword(x)$GUID
    StartTime <- keyword(x)$`$BTIM`
    EndTime <- keyword(x)$`$ETIM`
    CD45 Scaling <- keyword(x)$`$P43DISPLAY`
    CD45 Gain Setting <- keyword(x)$`$P43V`
    CD45 Bit Depth <- keyword(x)$`$P43B`

    Volume <- keyword(x)$`$VOL`
    Volume <- as.numeric(Volume)

    EventsInTheGate <- gs_pop_get_data(x, subset)
    Cells <- nrow(EventsInTheGate)[[1]] #[[1]] since the output is in a list

    Concentration <- (Cells*1000)/Volume
    Concentration <- Concentration*dilutionMultiplier

    Data <- data.frame(
        Date=Date,
        InstrumentSerial=InstrumentSerial,
        Specimen=Specimen,
        StartTime=StartTime,
        EndTime=EndTime,
        Volume=Volume, 
        Cells=Cells,
        Concentration=Concentration
        
        )

    return(Data)
}


```


:::{.callout-tip title="Problem 3"}
For CellConcentration, we retrieved both start and end time. Look up information on `lubridate` package, convert these times to a time-style format, and from acquired volume derrive uL/min at which each .fcs file was acquired. Did this vary at all across days?
:::