# Problem 1
Plugged in an external USB drive and trying to programmmatically create a folder and afterwards transfer some fcs-files to the newly created foilder.

Creating the folder on my USB (which was assigned the letter D)
```{r}
getwd()
list.files(path="D:/", full.names = TRUE, recursive = FALSE)
NewFolderLocation <- file.path("D:/", "Test")
dir.create(path=NewFolderLocation)
```

Copying fcs-files from data directory

```{r}
library(dplyr)
all_files <- list.files("data", full.names = TRUE, recursive = FALSE)
fcs_files <- list.files("data", pattern=".fcs", full.names = TRUE, recursive = FALSE)
file.copy(from=fcs_files, to=NewFolderLocation)
```

# Problem 2

Not sure I understand the question correctly, but if we want to list the files within a specific directory, we can use this code.

```{r}
#list.files(full.names = TRUE, recursive = TRUE)
#list.files(path="images", full.names = TRUE, recursive = TRUE)
Files <- list.dirs(path="data", full.names = TRUE, recursive = TRUE)
normalizePath(path=Files, winslash = "/", mustWork = TRUE)
```

# Problem 3

```{r}
system.file("extdata", package="FlowSOM")
system.file("extdata", package="flowWorkspace")
list.files(path="C:/Users/au160129/AppData/Local/Programs/R/R-4.5.2/library/FlowSOM/extdata", pattern=".fcs", full.names = TRUE, recursive = TRUE)
list.files(path="C:/Users/au160129/AppData/Local/Programs/R/R-4.5.2/library/flowWorkspace/extdata", pattern=".fcs", full.names = TRUE, recursive = TRUE)
```

It appears there is only one fcs file in the FlowSOM package and none in the flowWorkspace package.