Problem 1
Modify one of the simpler functions (SecondFunction or similar), provide your own argument names, and modify the message(), paste0 or print() functions to print a text style output. Generate a small vector of values, and iterate through your vector using one of the approaches we used.

```{r}
Bianca <- function(name, age){

  text <- paste0("Hi ", name, "! You are ", age, " years old.")

  print(text)

}

names <- c("Ana", "Carlos", "Julia", "Pedro")
ages <- c(22, 30, 25, 28)

for(i in 1:length(names)){

  Bianca(
    name = names[i],
    age = ages[i]
  )
}
Bianca(name=names, age=ages)
```

Problem 2
Using the initial framework for the CellConcentration function, retrieve several other keywords that are of interest, and incorporate them into the returned data.frame row.

```{r}
CellConcentration <- function(x, subset, dilutionMultiplier){

    Date <- keyword(x)$`$DATE`
    InstrumentSerial <- keyword(x)$`$CYTSN`
    TotalFileEvents <- keyword(x)$`$TOT`
    Specimen <- keyword(x)$GUID
    Operator <- keyword(x)$`$OP`
    StartTime <- keyword(x)$`$BTIM`
    EndTime <- keyword(x)$`$ETIM`

    Volume <- keyword(x)$`$VOL`
    Volume <- as.numeric(Volume)

    EventsInTheGate <- gs_pop_get_data(x, subset)
    Cells <- nrow(EventsInTheGate)[[1]]

    Concentration <- (Cells*1000)/Volume
    Concentration <- Concentration*dilutionMultiplier

    Data <- data.frame(
        Date=Date,
        InstrumentSerial=InstrumentSerial,
        Specimen=Specimen,
        Operator=Operator,
        StartTime=StartTime,
        EndTime=EndTime,
        Volume=Volume, 
        Cells=Cells,
        Concentration=Concentration
    )

    return(Data)
}
purrr::map(  .x=MyGatingSet[1],  .f=CellConcentration,  subset="CD19", dilutionMultiplier=100)
```
''
Problem 3
For CellConcentration, we retrieved both start and end time. Look up information on lubridate package, convert these times to a time-style format, and from acquired volume derrive uL/min at which each .fcs file was acquired. Did this vary at all across days?


```{r}
library(lubridate)
library(dplyr)

CellConcentration <- function(x, subset, dilutionMultiplier){

    Date <- keyword(x)$`$DATE`
    InstrumentSerial <- keyword(x)$`$CYTSN`
    TotalFileEvents <- keyword(x)$`$TOT`
    Specimen <- keyword(x)$GUID
    Operator <- keyword(x)$`$OP`

    StartTime <- keyword(x)$`$BTIM`
    EndTime <- keyword(x)$`$ETIM`

    Volume <- keyword(x)$`$VOL`
    Volume <- as.numeric(Volume)

    # converting time
    StartTime_hms <- hms(StartTime)
    EndTime_hms <- hms(EndTime)

    # difference
    AcquisitionTime <- as.numeric(
      difftime(EndTime_hms, StartTime_hms, units = "mins")
    )

    # uL/min
    FlowRate <- Volume / AcquisitionTime

    EventsInTheGate <- gs_pop_get_data(x, subset)
    Cells <- nrow(EventsInTheGate)[[1]]

    Concentration <- (Cells * 1000) / Volume
    Concentration <- Concentration * dilutionMultiplier

    Data <- data.frame(
        Date = Date,
        InstrumentSerial = InstrumentSerial,
        Operator = Operator,
        Specimen = Specimen,
        StartTime = StartTime,
        EndTime = EndTime,
        AcquisitionTime_Min = AcquisitionTime,
        Volume_uL = Volume,
        FlowRate_uL_min = FlowRate,
        Cells = Cells,
        Concentration = Concentration
    )

    return(Data)
}

Results <- purrr::map(  .x = MyGatingSet,  .f = CellConcentration,  subset = "CD19",
  dilutionMultiplier = 100)

Results <- bind_rows(Results)
Results

#The acquisition flow rate was highly consistent across acquisition days, remaining around 67–69 uL/min with only minor variation between files but some of them are negative and I don´t understand why
```