---
title: Week 1 homework
author: Aakash Modak
date: 16/03/2026
format: html
toc: true
---

# PeacoQC

PeacoQC is a package that performs quality control on FCS files - it can remove anomalies caused by clogging or variable flow rate that occurs when acquiring events from a flow cytometer. Its workflow requires assigning variables and working through step by step to result in cleaned flow data. FlowCore is a dependency package needed for it to function; it does xyz.

The basic workflow of the package begins with an initial step of assigning the path for your FCS file, and then reading the FCS file you assigned:
```{r}
#| eval: false
fileName <- system.file("extdata", "111.fcs", package="PeacoQC")
ff <- flowCore::read.FCS(fileName)
```

The next step is then defining the channels where margin/axes events are removed and which channels the QC should be performed. This step shoudl be performed before compensating and transforming any data:

```{r}
#| eval: false
channels <- c(1, 3, 5:14, 18, 21)
ff <- RemoveMargins (ff=ff, channels=channels,output="frame")
```

The next step is compensation and transformation of your data:

```{r}
#| eval: false
ff <- flowCore::compensate(ff, flowCore::keyword(ff)$SPILL)
```

The final step is then performing the QC and saving the results:

```{r}
#| eval: false
PeacoQC_res <- PeacoQC(ff, channels, determine_good_cells = "all", save_fcs = TRUE)
```

# Bioconductor Packages
