# Take-home Problems

"Problem 1"
Load a dataset into R, gate it however you like, and then export out a population of interest as their own .fcs files. Open them in either Floreada.io or the commercial software of your choice, and take a screenshot of how they look by two markers of interest. 

```{r}
PATH_TO_PAC_47fcs <-  file.path("course", "10_Downsampling", "homeworks", "biancakbeck")
PATH_TO_PAC_47fcs

PAC_47fcs <- list.files(PATH_TO_PAC_47fcs, pattern = ".fcs", full.names = TRUE)
PAC_47fcs
library(flowCore)

swab_47_f <- PAC_47fcs[1]
swab_47_f

read.FCS(filename = swab_47_f, transformation = FALSE, truncate_max_range = FALSE)

flowFrame <- read.FCS(filename=swab_47_f, transformation = FALSE, truncate_max_range = FALSE)


MFI_Matrix <- flowFrame@exprs
MFI_Matrix_Alternate <- exprs(flowFrame)
head(MFI_Matrix, 5)
ColumnNames <- colnames(MFI_Matrix)
ColumnNames
str(ColumnNames)
class(ColumnNames)
unname(ColumnNames)

#renaming the columns
colnames(MFI_Matrix )
colnames(MFI_Matrix)[5]
colnames(MFI_Matrix)[5] <- "SpikePositive"
colnames(MFI_Matrix)[6] <- "CD27Positive"
colnames(MFI_Matrix)[7] <- "IgMPositive"
colnames(MFI_Matrix)[8] <- "DUMP"
colnames(MFI_Matrix)[9] <- "CD19Positive"
colnames(MFI_Matrix)[10] <- "LIVE"
colnames(MFI_Matrix)[11] <- "CD138Positive"

colnames(MFI_Matrix)[13] <- "CD38Positive"

colnames(MFI_Matrix)[15] <- "IgAPositive"
colnames(MFI_Matrix)[16] <- "IgDPositive"
colnames(MFI_Matrix )

#ws
flowset <- read.flowSet(files = PAC_47fcs , truncate_max_range = FALSE, transformation = FALSE)
flowset

library(flowWorkspace)
cytoframe <- load_cytoframe_from_fcs(PAC_47fcs[1], truncate_max_range = FALSE, transformation = FALSE)

cytoframe

SwabFiles_F <- PAC_47fcs[stringr::str_detect(PAC_47fcs, "Fenotipagem_SWAB")]
SwabFiles_F

SFC_cytoset <- load_cytoset_from_fcs(SwabFiles_F, truncate_max_range = FALSE, transformation = FALSE)
SFC_GatingSet <- GatingSet(SFC_cytoset)
SFC_GatingSet
pData(SFC_GatingSet)



SFC_cytoset <- load_cytoset_from_fcs(SwabFiles_F, truncate_max_range = FALSE, transformation = FALSE)
SFC_GatingSet <- GatingSet(SFC_cytoset)
SFC_GatingSet
pData(SFC_GatingSet)

Metadata <- pData(SFC_GatingSet)
str(Metadata)

library(ggcyto)

#plotting
ggcyto(SFC_GatingSet[1], subset="root",
 aes(x="FSC-A", y="SSC-A")) + geom_hex(bins=100)
 str(SFC_GatingSet[1])
#para ver se ja tem transformações
gh_get_transformations(SFC_GatingSet[1])
colnames(SFC_GatingSet)
markernames(SFC_GatingSet)


#all
 ggcyto(SFC_GatingSet[1], subset="root",
 aes(x="FSC-A", y="SSC-A")) + geom_hex(bins=100)

#singlets
 ggcyto(SFC_GatingSet[1], subset="root",
 aes(x="FSC-A", y= "FSC-H")) + geom_hex(bins=100)

#live
 ggcyto(SFC_GatingSet[1], subset="root",
 aes(x="SSC-A", y= "BV510-A")) + geom_hex(bins=100)

#B cells
 ggcyto(SFC_GatingSet[1], subset="root",
 aes(x="BV421-A", y= "APC-Cy7-A")) + geom_hex(bins=100)

#Memory and Plasm
 ggcyto(SFC_GatingSet[1], subset="root",
 aes(x="BV421-A", y= "APC-A")) + geom_hex(bins=100)

#IgD-
 ggcyto(SFC_GatingSet[1], subset="root",
 aes(x="PE-Cy7-A", y= "SSC-A")) + geom_hex(bins=100)

#isotype
 ggcyto(SFC_GatingSet[1], subset="root",
 aes(x="PE-A", y= "A680-A")) + geom_hex(bins=100)

 #s+
  ggcyto(SFC_GatingSet[1], subset="root",
 aes(x="Alexa Fluor 488-A", y= "BV421-A")) + geom_hex(bins=100)

#to star transforming

  gh_get_transformations(SFC_GatingSet[[1]])

 Logicle <- logicle_trans()
str(Logicle)

Biexponential <- flowjo_biexp_trans()
str(Biexponential)

Asinh <- flowjo_fasinh_trans()
str(Asinh)

#to take only the fluorophores
SFC_Parameters <- colnames(SFC_GatingSet)
SFC_Parameters
FluorophoresOnly <- SFC_Parameters[!stringr::str_detect(SFC_Parameters, "FSC|SSC|Time")]
FluorophoresOnly

#transformer list
MyBiexTransform <- transformerList(FluorophoresOnly, Biexponential)
MyBiexTransform

#just to check and take the number of channel range, max value, pos, neg and widthbasis
str(MyBiexTransform[1:3])
str(MyBiexTransform$`APC-A`)

#taking out of my ram memory, making a gating set, in the function flowjo_biexp_trans putting the numbers that i got from str() function, this numbers that i will change to transform my data, and combining those with the transformer list that we want and reaplying to the gating set
 SFC_cytoset <- load_cytoset_from_fcs(PAC_47fcs,
 truncate_max_range = FALSE, transformation = FALSE)
SFC_GatingSet <- GatingSet(SFC_cytoset)
Biexponential <- flowjo_biexp_trans(
  channelRange = 4096,
  maxValue = 262144,
  pos = 3.8,
  neg = 0.5,
  widthBasis = -10)
MyBiexTransform <- transformerList(FluorophoresOnly, Biexponential)
transform(SFC_GatingSet, MyBiexTransform)

#B cells
 ggcyto(SFC_GatingSet[1], subset="root",
 aes(x="APC-Cy7-A", y= "BV421-A")) + geom_hex(bins=100)

#gating
library(remotes)
library(pak)
library(flowWorkspace)
library(ggplot2)
library(openCyto)
library(flowGate)
library(flowCore)

colnames(SFC_GatingSet)

GatingTable <- tibble::tribble(
  ~filterId,   ~dims,                         ~subset,

  "cells",     list("FSC-A", "SSC-A"),       "root",

  "singlets",  list("FSC-A", "FSC-H"),       "cells",

  "live",      list("SSC-A", "BV510-A"),     "singlets",

  "Bcells",    list("BV421-A", "APC-Cy7-A"), "live",

  "Memory",    list("BV421-A", "APC-A"),     "Bcells",

  "Plasma",    list("BV421-A", "APC-A"),     "Bcells"
)
GatingTable
SFC_cytoset <- load_cytoset_from_fcs(SwabFiles_F,
 truncate_max_range = FALSE, transformation = FALSE)
SFC_GatingSet <- GatingSet(SFC_cytoset)

SFC_Parameters <- colnames(SFC_GatingSet)
FluorophoresOnly <- SFC_Parameters[!stringr::str_detect(SFC_Parameters, "FSC|SSC|Time")]

Biexponential <- flowjo_biexp_trans(channelRange=4096, maxValue=262144,
     pos=4.5, neg=2, widthBasis=-500)
MyBiexTransform <- transformerList(FluorophoresOnly, Biexponential)
transform(SFC_GatingSet, MyBiexTransform)

gs_apply_gating_strategy(SFC_GatingSet, gating_strategy = GatingTable)

library(dplyr)
#' This function downsamples from a designated gate to our desired number
#' of cells, returning as a new .fcs file
#' 
#' @param x A GatingSet object, typically iterated in.
#' @param subset The gate from which to retrieve cell counts from 
#' @param inverse.transform Whether to revert values back to their
#' original untransformed values before export as an .fcs file, default
#' is set to TRUE
#' @param DownsampleCount The desired number of cells to downsample from
#' each gated population. If value is less than 1, subsets out the 
#' equivalent proportion from that specimen
#' @param addon An additional character value to add before .fcs in the GUID
#' keyword to tell the downsampled file apart from the original. 
#' @param StorageLocation A file.path to the folder you want to store the new downsampled
#' fcs file to. Default NULL results in .fcs file being stored in current working directory
#'  
#' 
 Downsampling <- function(x, subset, inverse.transform=TRUE, DownsampleCount,
  addon, StorageLocation=NULL, returnType="fcs"){
  EventsInTheGate <- gs_pop_get_data(x, subset,
   inverse.transform=inverse.transform)
  MeasurementData <- exprs(EventsInTheGate[[1]])
  MeasurementDataFramed <- as.data.frame(MeasurementData, check.names = FALSE)

  if (DownsampleCount < 1) {
      Count <- nrow(EventsInTheGate) # Original Count
      Count <- as.numeric(Count) #Sanity Check on Value Type
      Count <- Count*DownsampleCount # Target Cells
      Count <- round(Count, 0)
      DownsampleCount <- Count # Over-writting DownsampleCount used for downsampling
    }

  Downsampled_DataFrame <- slice_sample(MeasurementDataFramed, n = DownsampleCount, replace = FALSE)

  DownsampledMatrix <- as.matrix(Downsampled_DataFrame)

  flowFrame <- EventsInTheGate[[1, returnType = "flowFrame"]]
  OriginalParameters <- parameters(flowFrame)
  OriginalDescription <- keyword(flowFrame)

  OriginalName <- OriginalDescription$`GUID`
  UpdatedName <- paste0("_", addon, ".fcs")
  UpdatedGUID <- sub(".fcs", UpdatedName, OriginalName) #Swtiching out .fcs for Updated Name via the sub function
  OriginalDescription$`GUID` <- UpdatedGUID

  NewFCS <- new("flowFrame", exprs=DownsampledMatrix, parameters=OriginalParameters, description=OriginalDescription)

  if (is.null(StorageLocation)){StorageLocation <- getwd()}

  StoreFCSFileHere <- file.path(StorageLocation, UpdatedGUID)
  write.FCS(NewFCS, filename = StoreFCSFileHere, delimiter="#")

  return(StoreFCSFileHere)
}

library(purrr)

purrr::walk(.x=SFC_GatingSet, subset="cells", .f=Downsampling, DownsampleCount=2500, addon="cells",
  returnType="fcs", StorageLocation="course/10_Downsampling/homeworks/biancakbeck")

  #just made a screenshot and saved in the folder as screenshot1
```


Problem 2
In the example for `Downsampling()` we only changed one keyword (GUID), after substituting in our desired addon right before the .fcs. Since keyword use might vary by manufacturer, create a couple additional arguments for `Downsampling()` that allow you to change out the values for some additional keywords.

```{r}
#' This function downsamples from a designated gate to our desired number
#' of cells, returning as a new .fcs file
#' 
#' @param x A GatingSet object, typically iterated in.
#' @param subset The gate from which to retrieve cell counts from 
#' @param inverse.transform Whether to revert values back to their
#' original untransformed values before export as an .fcs file, default
#' is set to TRUE
#' @param DownsampleCount The desired number of cells to downsample from
#' each gated population. If value is less than 1, subsets out the 
#' equivalent proportion from that specimen
#' @param addon An additional character value to add before .fcs in the GUID
#' keyword to tell the downsampled file apart from the original. 
#' @param StorageLocation A file.path to the folder you want to store the new downsampled
#' fcs file to. Default NULL results in .fcs file being stored in current working directory
#'  
#' 
 Downsampling <- function(x, subset, inverse.transform=TRUE, DownsampleCount,
  addon, StorageLocation=NULL, returnType="fcs", NewTubeName = NULL){
  EventsInTheGate <- gs_pop_get_data(x, subset,
   inverse.transform=inverse.transform)
  MeasurementData <- exprs(EventsInTheGate[[1]])
  MeasurementDataFramed <- as.data.frame(MeasurementData, check.names = FALSE)

  if (DownsampleCount < 1) {
      Count <- nrow(EventsInTheGate) # Original Count
      Count <- as.numeric(Count) #Sanity Check on Value Type
      Count <- Count*DownsampleCount # Target Cells
      Count <- round(Count, 0)
      DownsampleCount <- Count # Over-writting DownsampleCount used for downsampling
    }

  Downsampled_DataFrame <- slice_sample(MeasurementDataFramed, n = DownsampleCount, replace = FALSE)

  DownsampledMatrix <- as.matrix(Downsampled_DataFrame)

  flowFrame <- EventsInTheGate[[1, returnType = "flowFrame"]]
  OriginalParameters <- parameters(flowFrame)
  OriginalDescription <- keyword(flowFrame)

  OriginalName <- OriginalDescription$`GUID`
  UpdatedName <- paste0("_", addon, ".fcs")
  UpdatedGUID <- sub(".fcs", UpdatedName, OriginalName) #Swtiching out .fcs for Updated Name via the sub function
  OriginalDescription$`GUID` <- UpdatedGUID
  if(!is.null(NewTubeName)){
  OriginalDescription$`$TUBENAME` <- NewTubeName
}

  NewFCS <- new("flowFrame", exprs=DownsampledMatrix, parameters=OriginalParameters, description=OriginalDescription)

  if (is.null(StorageLocation)){StorageLocation <- getwd()}

  StoreFCSFileHere <- file.path(StorageLocation, UpdatedGUID)
  write.FCS(NewFCS, filename = StoreFCSFileHere, delimiter="#")

  return(StoreFCSFileHere)
}

purrr::walk(.x=SFC_GatingSet, subset="cells", .f=Downsampling, DownsampleCount=2500, addon="cells",
  returnType="fcs", NewTubeName = "Downsampled", StorageLocation="course/10_Downsampling/homeworks/biancakbeck")

#cheking if it worked
  NewFile <- read.FCS(
  "course/10_Downsampling/homeworks/biancakbeck/Fenotipagem_SWAB_047_cells.fcs",
  transformation = FALSE
)

keyword(NewFile)$`$TUBENAME`
```

Problem 3"
Trickier - After concatenating out an .fcs file for a cell subset of your choice, reload it back into R, extract out both the exprs matrix, and the description list. Using the keywords that got added, figure out a way using dplyr to revert the numeric keys (denoted by "_key") in the exprs matrix back to their original character values as recorded in the keywords. 


```{r}
ConcatFCS <- read.FCS(  "course/10_Downsampling/homeworks/biancakbeck/Fenotipagem_SWAB_047_cells.fcs",  transformation = FALSE,  truncate_max_range = FALSE)

ExprsMatrix <- exprs(ConcatFCS)
ExprsMatrix
ExprsDF <- as.data.frame(ExprsMatrix,  check.names = FALSE)
ExprsDF
Keywords <- keyword(ConcatFCS)
Keywords
names(Keywords)[stringr::str_detect(  names(Keywords),
  "_key")]
names(keyword(ConcatFCS))

#None of the keywords ended with _key, I don't understand. If I had done this with the course data, would it have worked?
```