---
title: Solutions for week09
author: Jarkko Toivonen
date: today
format: 
    html:
        df-print: kable
        embed-resources: true
toc: true
toc-depth: 5
---

```{r}
#| label: setup
library(tibble)
library(purrr)
library(dplyr)
library(stringr)
library(ggcyto)
```

# Problem 1

> Modify one of the simpler functions (SecondFunction or similar), provide your own argument names, and modify the message(), paste0 or print() functions to print a text style output. Generate a small vector of values, and iterate through your vector using one of the approaches we used.

```{r}
f <- function(name, level, max = 5) {
  v <- c(rep(" ", max - level), rep("*", level))
  prefix <- v |> paste(collapse = "")
  suffix <- v |> rev() |> paste(collapse = "")
  sprintf("%s %s %s", prefix, name, suffix)
}
1:3 |> map_chr(\(i) f("John", i)) %>% cat(sep = "\n")
```

# Problem 2

> Using the initial framework for the CellConcentration function, retrieve several other keywords that are of interest, and incorporate them into the returned data.frame row.

Read the files and create a gating set.

```{r}
#| label: Load data
data_path <- file.path("data") # Quarto Rendering path

fcs_files <- list.files(data_path, pattern=".fcs", full.names=TRUE)
MyCytoset <- load_cytoset_from_fcs(fcs_files, transformation=FALSE, truncate_max_range=FALSE)
MyGatingSet <- GatingSet(MyCytoset)
```

Check all the keywords.

```{r}
#| label: Keywords
kws <- keyword(MyGatingSet[[1]])
head(kws)
```

```{r}
#| label: CellConcentration
CellConcentration <- function(x){
    kw <- keyword(x)
    Date <- kw$`$DATE` # Day-Month-Year
    InstrumentSerial <- kw$`$CYTSN`
    Creator <- kw$CREATOR             # Added
    TotalFileEvents <- kw$`$TOT`
    Specimen <- kw$GUID
    FCSversion <- kw$FCSversion       # Added
    StartTime <- kw$`$BTIM`
    EndTime <- kw$`$ETIM`
    Volume <- kw$`$VOL`

    # Count the number of lasers
    Lasers <- names(kw) |> keep(\(x) str_detect(x, "LASER[0-9]+NAME")) %>% length()

    res <- tibble(
        Date = Date,
        InstrumentSerial = InstrumentSerial,
        Creator = Creator,
        TotalFileEvents = TotalFileEvents,
        Specimen = Specimen,
        FCSversion = FCSversion,
        StartTime = StartTime,
        EndTime = EndTime,
        Volume = Volume,
        Lasers = Lasers
    )
    res
}
CellConcentration(MyGatingSet[[1]])
```

# Problem 3

> For CellConcentration, we retrieved both start and end time. Look up information on lubridate package, convert these times to a time-style format, and from acquired volume derrive uL/min at which each .fcs file was acquired. Did this vary at all across days?

Data File Standard for Flow Cytometry, Version FCS 3.1
https://pmc.ncbi.nlm.nih.gov/articles/PMC2892967/

> Sample Volume: A standardized way to communicate the sample volume
has been introduced. The value of a new optional $VOL keywords
is a floating point number expressing the sample volume in nanoliters.

But based on the scale, about 30, it looks that the volume in the file could be 
in microlitres (uL) instead!

I will assume that the unit is microlitres.

```{r}
#| label: CellConcentration3
#' This function will figure out the cell concentration
#'  for a given gated population in an .fcs file
#'
#' @param x A GatingSet object (typically iterated in)
#' @param subset The gate from which to retrieve cell counts from
#' @param dilutionMultiplier Multiplies the estimated concentration
#'  to bring correct digits. 
#' 
CellConcentration3 <- function(x, subset, dilutionMultiplier){
    Date <- keyword(x)$`$DATE` # Day-Month-Year
    InstrumentSerial <- keyword(x)$`$CYTSN`
    TotalFileEvents <- keyword(x)$`$TOT`
    Specimen <- keyword(x)$GUID
    StartTime <- keyword(x)$`$BTIM`
    EndTime <- keyword(x)$`$ETIM`

    StartTime <- hms::as_hms(StartTime)
    EndTime <- hms::as_hms(EndTime)
    Duration <- EndTime - StartTime

    Volume <- keyword(x)$`$VOL`
    Volume <- as.numeric(Volume)    # In nanolitres or microlitres, maybe the latter

    EventsInTheGate <- gs_pop_get_data(x, subset)
    Cells <- nrow(EventsInTheGate)[[1]] # [[1]] since the output is in a list

    Concentration <- (Cells*1000)/Volume   # Apparently this converts to cells per millilitres?
    Concentration <- Concentration*dilutionMultiplier

    CellsPerSecond <- Cells / as.numeric(Duration, unit = "secs")

    Flow <- Volume / as.numeric(Duration, unit = "mins")  # microlitres per minute (uL/min)
    FlowUnit <- "uL/min"

    Data <- tibble(
        Date=Date,
        InstrumentSerial=InstrumentSerial,
        Specimen=Specimen,
        StartTime=StartTime,
        EndTime=EndTime,
        Duration = Duration,
        Volume=Volume,    # In nanolitres or microlitres, maybe the latter
        Cells=Cells,
        Concentration=Concentration,
        CellsPerSecond = CellsPerSecond,
        Flow = Flow,
        FlowUnit = FlowUnit
        )

    return(Data)
}
```


The flow is of the diluted liquid.

Transform the data.

```{r}
#| label: Transform data
MainDetectorsOnly <- c("V3-A", "B4-A", "R1-A")

Biexponential <- flowjo_biexp_trans(channelRange=4096, maxValue=262144,
     pos=4.5, neg=2, widthBasis=-500)
MyBiexTransform <- transformerList(MainDetectorsOnly, Biexponential)
transform(MyGatingSet, MyBiexTransform)
```

Gate it.

```{r}
#| label: Gate table
GatingTable <- tibble::tribble(
  ~filterId,    ~dims,                          ~subset,           
  "CD45",       list("B4-A", "SSC-A"),            "root",     
  "CD19",       list("V3-A", "SSC-A"),            "CD45",   
  "CD14",       list("R1-A", "SSC-A"),            "CD45", 
  "TandNKcells",  list("V3-A", "SSC-A"),            "CD45",
)
GatingTable
```


```{r}
#| label: Create gates
filename <- file.path(data_path, "gated.gs")
if (! file.exists(filename)) {
    flowGate::gs_apply_gating_strategy(MyGatingSet, gating_strategy = GatingTable)
    save_gs(MyGatingSet, filename)
} else {
    MyGatingSet <- load_gs(filename)
}
```

```{r}
#| label: Show gate hierarchy
plot(MyGatingSet)
```

```{r}
#| label: Visualize gates
ggcyto::autoplot(MyGatingSet[[6]]) + theme_bw()
```




```{r}
#| label: Result table
population <- "root"
get_results_for_population <- function(gating_set, population) {
    df <- gating_set |>
        map(\(x) CellConcentration3(x, subset=population, dilutionMultiplier=100)) |>
        bind_rows()
    df
}
df <- get_results_for_population(MyGatingSet, population)
df
```


```{r}
#| label: Result plot of flow rate
df %>% 
    ggplot(aes(Specimen, Flow)) +
    geom_point() +
    lims(y = c(0, NA)) +
    labs(x = "Specimen", y = "Flow rate [uL/min]",
    title = "Flow rate", subtitle = "Root population") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The flow rate is almost constant across specimen so the day doesn't affect it.

Let's also check the population-wise cell rate, so that the creation of 
gates is not totally wasted.


```{r}
populations <- gs_get_pop_paths(MyGatingSet)
populations
```

Get results for all populations.

```{r}
#| label: Process all gates
dfs <- populations |> 
    set_names() |>     # Results in a named list, where the name is the same as the value
    map(\(population) get_results_for_population(MyGatingSet, population))
all <- bind_rows(dfs, .id = "population")   # The name of the list element goes to the population column
all |> head()
```

```{r}
#| label: Result plot of cell rate
all %>% 
    ggplot(aes(Specimen, CellsPerSecond, color = population, group = population)) +
    geom_point() +
    geom_line() +
    lims(y = c(0, NA)) +
    labs(x = "Specimen", y = "Cell rate [cells/s]",
    title = "Flow rate", subtitle = "Root population") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Hmm, the amount of rubbish seems to vary.