### Importing data from week 6 lesson

```{r}
library(dplyr)
library(ggplot2)
library(ggbeeswarm)
StorageLocation <- file.path("Cytometry Carlos", "06_Visualizing", "data") # When interactively writing the code 
TheCSV <- list.files(StorageLocation, pattern=".csv", full.names=TRUE)
Data <- read.csv(TheCSV, check.names=FALSE)
Data <- Data |> 
    mutate(TcellProportion=Tcells_count/CD45_count) |>
    mutate(TcellFrequency=TcellProportion *100) |>
    mutate(TcellFrequency=round(TcellFrequency, 1))
```

### Problem 1
In this session, we created beeswarm-style boxplot to display our T-cell frequencies on the y-axis, and timepoint on the x-axis. Using the concepts covered this week, swap out “timepoint” for the “Condition” variable. Adjust other layer arguments accordingly until you can return a similar plot at the end of the class. Finally, figure out how to switch around the order the Condition values are displayed on the x-axis.
```{r}
str(Data) #Condition is a character value
Data$Condition <- factor(Data$Condition) #We transform it into a factor

shape_sex <- c("Female" = 22, "Male" = 21)
fill_sex <- c("Female" = "white", "Male" = "black") #Creating the aesthetics list as in the index file
BSPlot <- ggplot(Data, aes(x = Condition, y = TcellFrequency)) + 
    geom_boxplot() + 
    geom_beeswarm(size=2.5, cex=2.5, aes(shape = infant_sex, fill = infant_sex)) +
    scale_shape_manual(values = shape_sex) +  scale_fill_manual(values = fill_sex)
BSPlot #Building and printing the plot as in the index file
```

To switch the order of Condition, a factor, we need to have a look at the arguments inside the factor function:
```{r}
?base::factor
```

To change the order, we can define it using the levels argument, and then set the ordered argument to true:
```{r}
Data$Condition <- factor(Data$Condition, levels = c("Ctrl", "SEB", "PPD"), ordered = T) #We set the new order
OrderedPlot <- ggplot(Data, aes(x = Condition, y = TcellFrequency)) + 
    geom_boxplot() + 
    geom_beeswarm(size=2.5, cex=2.5, aes(shape = infant_sex, fill = infant_sex)) +
    scale_shape_manual(values = shape_sex) +  scale_fill_manual(values = fill_sex) 
OrderedPlot
```

Additionally, we could define the order inside the aesthetics of gglot
```{r}
OrderedPlot2 <- ggplot(Data, aes(x = factor(Condition, levels = c("Ctrl", "SEB", "PPD"), ordered = T), y = TcellFrequency)) + 
    geom_boxplot() + 
    geom_beeswarm(size=2.5, cex=2.5, aes(shape = infant_sex, fill = infant_sex)) +
    scale_shape_manual(values = shape_sex) +  scale_fill_manual(values = fill_sex) 
OrderedPlot2
```

### Problem 2
Circle back to the CytoML-ggcyto flowplot, and modify it until happy with the visual appearance. You may use any resource on the internet to assist, but you must document your steps so that we can also repeat them.
```{r}
library(CytoML)
library(ggcyto)
FlowJoWsp <- list.files(path = StorageLocation, pattern = ".wsp", full = TRUE)
ws <- open_flowjo_xml(FlowJoWsp)
gs <- flowjo_to_gatingset(ws=ws, name=1, path = StorageLocation, additional.keys = "GROUPNAME")
Plot <- ggcyto(gs[6], subset="Tcells", aes(x="CD8", y="CD4")) + geom_hex(bins=100)
Plot
```

Here I'll try to use my previous knowledge of ggplot2 (basically the labs and theme functions):

```{r}
myPlot <- ggcyto(gs[6], subset="Tcells", aes(x="CD8", y="CD4")) + 
    geom_hex(bins=128) + #I believe a power of 2 looks better for some reason (?)
    theme_classic() +
    labs(title = "T cells", x = "CD8 - BV650", y = "CD4 - BUV805") + #labs() function allows to change the text on the label
    theme(plot.title = element_text(hjust = 0.5, size = 20, face = "bold"), #changing the plot title to a bigger and bold font and centering it
            axis.title = element_text(size = 15), #increasing font size of the axis labels
            axis.text = element_text(size = 10), #increasing font size of the axis numbers
            rect = element_blank(), strip.text = element_blank(), #removing the sample name and the rectangle
            legend.position = "none" #removing the legend
    )

myPlot
```

I will also try to draw a couple of lines to simulate a quad gate (I really don't know if it's going to work while I'm writing this)

```{r}
quadPlot <- myPlot +
    geom_line(aes(x = 120)) +
    geom_line(aes(y = 120))
quadPlot
```

More or less worked! I have not been able to make them reach the axis though.

### Problem 3
In Mismatched Assumptions, we saw two examples of a histogram/density overlay showing the distribution of a variable on the x-axis. Similar to what we did during class to show values according to a different data column, try to modify the plot to show data on the basis of group (whether condition, ptype, infant_sex, etc.) similar to what you can see here.

```{r}
dPlot <- ggplot(Data, aes(TcellFrequency)) #We assign the Data and the aesthetic layers to an object
dPlot + geom_density() #A normal density plot would look like this
```

Now, if we want to show the data using different colors for each condition, we need to map this inside the density layer

```{r}
dPlot + geom_density(aes(colour = Condition))
dPlot + geom_density(aes(fill = Condition))
```

This works with both colour and fill aesthetics arguments, depending on how we want to do this. Now, we can combine both using our color's choice and setting the alpha argument so that overlapped lines are actually visible

```{r}
color_condition <- c("Ctrl" = "#00AAFF", "SEB" = "#AAFF00", "PPD" = "#FF00AA") #We create our custom color list

dPlot + geom_density(aes(colour = Condition, fill = Condition), alpha = 0.3) + 
    scale_color_manual(values = color_condition) +
    scale_fill_manual(values = color_condition)
```

And that would do it! All condition groups seem to have the same distribution. We can use any other grouping variable:

```{r}
color_sex <- c("Male" = "#00AAAA", "Female" = "#AAAA00") #We create our custom color list

dPlot + geom_density(aes(colour = infant_sex, fill = infant_sex), alpha = 0.3) + 
    scale_color_manual(values = color_sex) +
    scale_fill_manual(values = color_sex)
```