---
title: "Take -home problems Week5"
author: "Denise Brito"
format: html
execute: 
    code-fold: show
    warning: false
---

## Problem 1 : Mutate GatingSet
### 1.1 GatingSet

```{r}
#Libraries needed
library(stringr)
library(CytoML)
library(dplyr)
library(ggplot2)
library(ggcyto)
```

```{r}
#Path and Files
Folder <- file.path("data")
FlowJoWsp <- list.files(path=Folder , pattern="wsp", full=TRUE)
FlowJoWsp
```

```{r}
#Workspace Selection
ThisWorkspace <- FlowJoWsp[stringr:: str_detect(FlowJoWsp, "Opened")]
ThisWorkspace
ws <- open_flowjo_xml(ThisWorkspace)
```

```{r}
#Gating Set
gs <- flowjo_to_gatingset(ws=ws, name=1, path=Folder, additional.keys = "GROUPNAME")
```

### 1.2 Retrieve the data.frame from cell counts per gate

```{r}
plot(gs)
PopCount <- gs_pop_get_count_fast(gs)
head(PopCount,5)
```

### 1.3 Mutate a new column showing percent of the parent gate. 

```{r}
PopCount_mutated <- PopCount |> mutate(percent_of_parent_gate = round((Count / ParentCount)*100,2))
head(PopCount_mutated,5)
```


## Problem 2 : Removing FCS files and Documenting Errors

**I deleted: 2025_07_26_AB_02_INF179_00_Ctrl.fcs**

Error detected here:

gs <- flowjo_to_gatingset(ws=ws, name=1, path=Folder, additional.keys = "GROUPNAME")

"FCS not found for sample 00_Ctrl.fcs_INF179 from searching the file extension: .fcs"

**Then I deleted all files except: 2025_07_26_AB_02_INF052_00_SEB.fcs**

Error detected :

FCS not found for sample 00_Ctrl.fcs_INF052 from searching the file extension: .fcs
FCS not found for sample 00_Ctrl.fcs_INF100 from searching the file extension: .fcs
FCS not found for sample 00_Ctrl.fcs_INF179 from searching the file extension: .fcs

**Finally, deleted all samples:**

error : ! No samples in this workspace to parse!
    ▆
 1. └─CytoML::flowjo_to_gatingset(...) at Cytometry%20in%20R/CytometryInR-Week5/take-home-problems-week5.qmd:38:1
 2.   ├─base::do.call(parse_workspace, args)
 3.   └─CytoML (local) `<fn>`(...)

## Problem 3 : Visualize TNFa vs IFNg for the various cell populations,  across both Ctrl and SEB samples

For this exercise i used the ggcyto vignette documentation:
https://www.bioconductor.org/packages/release/bioc/vignettes/ggcyto/inst/doc/Top_features_of_ggcyto.html#low_level:_ggplot

### 3.1 Discover TNF and IFN parameters name:

```{r}
fr <- gh_pop_get_data(gs[[1]]) #to extract data from file 1
pData(parameters(fr)) # to identify the parameter names for TNFa and IFNg
```

### 3.2 Plot tests
### 3.2.1 Plot test 1 - Sample 1 ,  CD4+ cells, bins=100

TNFa -> Comp-PE-Dazzle594-A       
IFNg -> Comp-BV750-A

```{r}
ggcyto(gs[4], subset="CD4+", aes(x="Comp-PE-Dazzle594-A", y="Comp-BV750-A")) + geom_hex(bins=100)
```

### 3.2.2 Plot test 2 - Sample 1 ,  CD4+ cells, bins=1000

```{r}
ggcyto(gs[4], subset="CD4+", aes(x="Comp-PE-Dazzle594-A", y="Comp-BV750-A")) + geom_hex(bins=1000)
```

### 3.2.3 Plot test 2 - Sample 1 ,  CD4+ cells, bins=10

```{r}
ggcyto(gs[4], subset="CD4+", aes(x="Comp-PE-Dazzle594-A", y="Comp-BV750-A")) + geom_hex(bins=10)
```

### 3.2.4 Plot test 2 - Sample 1 ,  CD4+ cells, bins=80

```{r}
ggcyto(gs[4], subset="CD4+", aes(x="Comp-PE-Dazzle594-A", y="Comp-BV750-A")) + geom_hex(bins=80)
```


### 3.3 Ctrl and SEB samples

Chosen populations:
CD4+
CD8+
CD4-CD8-

1st sample
```{r}
#Layout for sample 1
sample_1 <- gs[[1]]

gs_get_pop_paths(sample_1, path= "auto") # To check the indexes of the populations I'm interested in.

pops_to_plot<- gs_get_pop_paths(sample_1, path= "auto")[c(6:20)]
pops_to_plot 

#To create several plots
plots_for_sample1 <- autoplot(sample_1, pops_to_plot, bins= 80)
plots_for_sample1

#To view plots in one row
gt <- ggcyto_arrange(plots_for_sample1, nrow=1) 
plot(gt)
```

2nd Sample
```{r}
#Layout for sample 2
sample_2 <- gs[[4]]

gs_get_pop_paths(sample_2, path= "auto") # To verify the indexes of the populations I'm interested in.

pops_to_plot_2<- gs_get_pop_paths(sample_2, path= "auto")[c(6:20)]
pops_to_plot_2 

#To create several plots
plots_for_sample2 <- autoplot(sample_2, pops_to_plot_2, bins= 80)
plots_for_sample2

#To view plots in 1 row
gt2 <- ggcyto_arrange(plots_for_sample2, nrow=1) 
plot(gt2)
```

```{r}
gt3 <- gridExtra::gtable_rbind(gt,gt2)
plot(gt3)
```

