---
title: "Homework week 9"
author: "Carlos Cuño"
date: 04-17-2026
format: html
---


## Problem 1
### Modify one of the simpler functions (SecondFunction or similar), provide your own argument names, and modify the message(), paste0 or print() functions to print a text style output. Generate a small vector of values, and iterate through your vector using one of the approaches we used.

```{r}
#Importing libraries
library(flowWorkspace)
library(purrr)
```

I start by creating the function and seeing that it works with the default values. I also added the roxygen2 skeleton, which was actually quite cool to learn because it seems that it's the first step to create the documentation for a function/package.
```{r}
#' This function compares two given fluorophores
#' 
#' @param Fluorophore A character object containing a good fluorophore choice
#' @param WorseFluorophore A character object containing a bad fluorophore choice
#' 
MyOwnFunction <- function(Fluorophore = "RB613", WorseFluorophore = "PerCP-Cy5.5"){ #My own opinion was added as default values for both arguments
    output <- paste0("Without a doubt, ", Fluorophore, " is a better option than ", WorseFluorophore)
    print(output)
}
MyOwnFunction()
```

Good, now we just need to create the vector and iterate the function through it
```{r}
good.fluorophore.list <- c("BUV805","BV711", "PE", "RB613")
map(good.fluorophore.list, MyOwnFunction)
```

It works, but now I wonder: can I iterate the function through a vector that corresponds to the second argument of the function?
```{r}
bad.fluorophore.list <- c("BUV395", "FITC", "PE-Cy7", "APC-Cy7")
map(bad.fluorophore.list, MyOwnFunction(WorseFluorophore = bad.fluorophore.list))
```

I came up with this solution. It passes the list to the second arguments and returns the desired output, but then returns a list of NULL values, so there is probably a better way to do this.

## Problem 2
### Using the initial framework for the CellConcentration function, retrieve several other keywords that are of interest, and incorporate them into the returned data.frame row.

I first import a GatingSet from my week 8 homework:
```{r}
#gs.location <- file.path("Cytometry Carlos", "09_Functions", "homeworks", "carloscgomiz", "Manual Gating set.gs")
gs.location <- file.path("Manual Gating set.gs")
gs <- load_gs(gs.location)
plot(gs)
```

Next, I want to see the list of keywords in these files to decide which of them can be of interest. To do that, we extract all keywords from a sample and then create a vector with all the keyword names:
```{r}
keywords.list <- keyword(gs[[1]])
keywords <- names(keywords.list)
```

Now, I will create a function that saves all the data regarding cytometer manipulation: who acquired the sample, when, where and how:
```{r}
ExtractKeywords <- function(x){
    Sample <- keyword(x)$`$FIL`
    Cytometer <- keyword(x)$`$CYT`
    CytometerSN <- keyword(x)$`$CYTSN`
    Operator <- keyword(x)$`$OP`
    BeginTime <- keyword(x)$`$BTIM`
    EndTime <- keyword(x)$`$ETIM`
    Volume <- keyword(x)$`$VOL`
    FlowRate <- keyword(x)$`$FLOWRATE`

    dataframe <- data.frame(
        Sample=Sample,
        Cytometer=Cytometer,
        CytometerSN =CytometerSN,
        Operator=Operator,
        BeginTime=BeginTime,
        EndTime=EndTime,
        Volume=Volume,
        FlowRate=FlowRate
    )
    return(dataframe)
}
```

Does the function work on a single .fcs file?
```{r}
ExtractKeywords(gs[[1]])
```

Yes! Finally, I'll try to iterate it through both .fcs samples:

```{r}
map(gs, ExtractKeywords)
```

Good. Finally, I extract these keywords to a dataframe:

```{r}
DataFrame <- map(gs, ExtractKeywords) |> dplyr::bind_rows()
```

## Problem 3
### For CellConcentration, we retrieved both start and end time. Look up information on lubridate package, convert these times to a time-style format, and from acquired volume derrive uL/min at which each .fcs file was acquired. Did this vary at all across days?
In this case, I will import the GatingSet and a summarized version of the function from the course:

```{r}
#course.gs <- load_gs(file.path("Cytometry Carlos", "09_Functions", "data", "TheGatingSet.gs"))
course.gs <- load_gs("TheGatingSet.gs")

SummarizedFunction <- function(x){
    Date <- keyword(x)$`$DATE` # Day-Month-Year
    InstrumentSerial <- keyword(x)$`$CYTSN`
    TotalFileEvents <- keyword(x)$`$TOT`
    Specimen <- keyword(x)$GUID
    StartTime <- keyword(x)$`$BTIM`
    EndTime <- keyword(x)$`$ETIM`

    Volume <- keyword(x)$`$VOL`
    Volume <- as.numeric(Volume)

    Data <- data.frame(
        Date=Date,
        InstrumentSerial=InstrumentSerial,
        Specimen=Specimen,
        StartTime=StartTime,
        EndTime=EndTime,
        Volume=Volume
        )

    return(Data)
}
```

Now I generate the dataframe:

```{r}
DataFrame <- map(course.gs, SummarizedFunction) |> dplyr::bind_rows()
class(DataFrame$StartTime)
class(DataFrame$EndTime)
```

The times are in HH:MM:SS format, but stored as characters. Let's load the lubridate package and see what I can do:

```{r}
library(lubridate)
?lubridate::hms #This function might do it
DataFrame$StartTime = hms(DataFrame$StartTime)
DataFrame$EndTime = hms(DataFrame$EndTime)
class(DataFrame$StartTime)
class(DataFrame$EndTime)
```

Now they have been transformed to "Period" class. Can I now substract one from the other?
```{r}
DataFrame = DataFrame |>
    dplyr::mutate(RunTime = EndTime - StartTime)
DataFrame$RunTime
```

Yes! It has given me the elapsed time for each sample. However, I would like to transform it into numeric values in order to facilitate the calculation:

```{r}
DataFrame$RunTime = as.numeric(DataFrame$RunTime)
```

Good. Finally, I'll calculate the flow rate in uL/min (creating a small function![not necessary but it's good to practice]):

```{r}
flowrate <- function(volume, time){
 flowrate = volume * 60/time
 return(flowrate)
}

DataFrame = DataFrame |> dplyr::mutate(FlowRate = flowrate(volume=Volume, time = RunTime))
DataFrame$FlowRate
```